Title here

### 2016-09-26 portfolio/text/ms.Rmd
Incorporate Ole's edits

---
output: html_document
---
```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
```
```{r get-dat, cache=FALSE}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
```

<!-- Ole likes the first or fourth title best, "
Benefit of diversification are unavailable to many fishermen
" -->

# Ole Abstract

<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->

A fundamental challenge for individuals who rely on the extraction of natural resource for their livelihood is income variability.  
Fishers, loggers, farmers, miners, and other  all face uncertainty in income driven by a mix of environmental, biological, and economic (e.g. price) varibility.
A key component underlying the sustainability of these industries is identifying both regulatory actions and individual behavior affect income variability and by extension the economic stability and sustainable use of natural resources.

For commercial fisheries, recent work has identified that targeting a diversity of species can reduce income variability.
However, any diversity-variability relationship is embedded in a regulatory framework that restricts individual ability to unilaterally change fishing behavior.  
Here we identify three primary axes by which fishers in Alaska, one of the world's most productive fishery regions, can potentailly diversify fishing activities and understand how individual actions along each diversity axis affect both revenue and revenue variability. 
We construct a novel statistical model to simultaneously assess the value to individuals fishers of choosing to 1) own permits that allow fishers that target a diverse fish species assemblage, 2) diversify by owning multiple permits, or 3) catch a greater species within the constraints of a given permit.

Thus, we present a unique approach to understanding important aspects of diversity relevant to individaul fishers' decisions. 
We show contrastings effects of diversity on revenue and revenue variability along the three axes. First, we show (results for choosing a permit). Next we show (results for owning multiple permits).  Finally we show (results for individual permit diversification).

Thus, while there are clear benefits to some flavors of diversification, others don't do much.  Importantly, the most effective measure for increasing revenue and decreasing risk - purchasing additional permits to participate in additional fisheries - was also the most costly

We find that
fishing an additional permit is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits);
fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity);
but that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

# Original Abstract

Individuals relying on natural resources as their primary source of income
often experience high levels of income variability
due to the combined effects of environmental, biological, and price variation.
Commercial fishers are a prime example and
recent work has sought to identify regulations and fishing behaviours
that can reduce this variability.
One candidate has been that fishing higher levels of species diversity
decreases income variability.
However, it is unknown if this species-diversity effect
on income variability holds
across multiple approaches to diversifying catch,
at the level of individual permit holders,
and across different fishing strategies.
Here, we simultaneously estimate the effects of three
methods of diversification for individual participants in Alaskan fisheries
on revenue and revenue variability:
(1) diversifying by owning multiple permits,
(2) diversifying by owning permits that inherently increase species diversity,
and (3) increasing species diversity within the constraints of a permit set.
We control for changes in effort and other
latent effects through time.
We find that
fishing an additional permit is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits);
fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity);
but that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

@@

- Farming: crop rotation has been proposed as metric for buffering against short term
  variation and long term trends (like climate change)
- Logging: Maintaining a diverse mix of tree species, age structure, and space to buffer against things like wildfires and price varition. 
- but in some situations, external forces (prices, demand) may discourage the
  adoption of these diverse strategies and 
  help promote specialization [@bradshaw2004]

@@

We develop a novel statistical approach for understanding the effect of diveristy on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community levels [@sethi2014]. First, we  summarize trends in diversification across fishing fleets in Alaska. Next, we identified three main axes by which fishers may diversify: 1), 2), 3). Finally, we constructed a model to simultaneously estimate the relationship between each of these three axes of diversity and both annual revenue and risk for individual fishers over nearly ZZ years. Our hierachical model makes use of entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits).  It is the first model that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.

strategies where both specialization and diversification are extremely profitable. There's a number of fishing permits that constrain the potential diversity of species landed (e.g. herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.* 
*perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*

@@

-Ole: I'd start with a little reminder of what the purpose of the model is for. Like:
To jointly model revenue and revenue variability as a function of individual-level covariates, we use a variance function regression model, which separately models change in the mean and variance as
a function of covariates [e.g. @davidian1987; @hall1989]. Where traditional
linear models or hierarchical linear models generally assume constant
variance [across what? covariates?], the variance function regression model allows variance to be
heteroskedastic. For application to economic data on individual fishing
portfolios, this type of model allows certain strategies and levels of
diversification to be more or less variable than others.

@@

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$ 
(Supporting Information, Table \ref{tab:strategies})
and expected $\log(R_{i,t})$ is modelled as:
OLE: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid 
potential confusion like using the term “expected log(R)”.  I also just don't like hats... but that's my problem.

>>>>

```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
```
```{r get-dat, cache=FALSE}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
```

A fundamental challenge for individuals who rely on the extraction of natural
resource for their livelihood is income variability.
Fishers, loggers, farmers, miners, and others
all face uncertainty in income driven by a mix of
environmental, biological, and economic (e.g. price) variability.
A key component underlying the sustainability 
of these industries is identifying both
regulatory actions and individual behaviour affect income variability 
and by extension the economic stability and 
sustainable use of natural resources.   
For commercial fisheries, recent work has identified that 
targeting a diversity of species can reduce income variability.
However, any diversity-variability relationship 
is embedded in a regulatory framework that restricts individual ability 
to unilaterally change fishing behaviour.  
Here we identify three primary axes by which fishers in Alaska, 
one of the world's most productive fishery regions, can 
potentially diversify fishing activities and understand how 
individual actions along each diversity axis affect 
both revenue and revenue variability. 
We construct a novel statistical model to simultaneously 
assess the value to individuals fishers of choosing to 
(1) own permits that allow fishers that target a diverse fish species assemblage, 
(2) diversify by owning multiple permits, or 
(3) catch a greater species within the constraints of a given permit.  
Thus, we present a unique approach to understanding important aspects of
diversity relevant to individual fishers' decisions. 
We show contrasting effects of diversity on 
revenue and revenue variability along the three axes. 
First, we show that fishing permits with 
higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity).
Next, we show that fishing an additional permit 
is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits);
Finally, we show that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue the following year.

Thus, while there are clear benefits 
to some flavours of diversification, others don't do much.
Importantly, the most effective measure for 
increasing revenue and decreasing risk --- 
purchasing additional permits to participate in additional fisheries --- 
was also the most costly.

At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

@@

- Farming: crop rotation has been proposed as metric for buffering against short term
  variation and long term trends (like climate change)
- Logging: Maintaining a diverse mix of tree species, age structure, and space
  to buffer against things like wildfires and price variation. 
- but in some situations, external forces (prices, demand) may discourage the
  adoption of these diverse strategies and 
  help promote specialization [@bradshaw2004]

@@

We develop a novel statistical approach for understanding the effect of
diversity on both revenue and risk that builds on results from the vessel
[@kasperski2013], fleet [@sethi2012b], and community levels [@sethi2014].
First, we  summarize trends in diversification across fishing fleets in Alaska.
Next, we identified three main axes by which fishers may diversify: 1), 2), 3).
Finally, we constructed a model to simultaneously estimate the relationship
between each of these three axes of diversity and both annual revenue and risk
for individual fishers over nearly ZZ years. 
Our hierarchical model makes use of
entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits).  
It is the first model that jointly explores the effect of diversity on revenue
and risk at the individual level and the first considers the multiple axes of
diversity simultaneously.

strategies where both specialization and diversification are extremely
profitable. There's a number of fishing permits that constrain the potential
diversity of species landed (e.g. herring, halibut, sablefish, or crab permits
that allow the user to only land that species) versus broader 'finfish' permits
that allow users to land a wide variety of groundfish spp.*

*perhaps we also need to have a sentence or two on the implicit constraints of
geography and gear / vessel size on species diversification*

@@

-Ole: I'd start with a little reminder of what the purpose of the model is for. Like:
To jointly model revenue and revenue variability as a function of
individual-level covariates, we use a variance function regression model, which
separately models change in the mean and variance as
a function of covariates [e.g. @davidian1987; @hall1989]. Where traditional
linear models or hierarchical linear models generally assume constant
variance [Ole: across what? covariates?], the variance function regression model allows variance to be
heteroskedastic. For application to economic data on individual fishing
portfolios, this type of model allows certain strategies and levels of
diversification to be more or less variable than others.

@@

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$ 
(Supporting Information, Table \ref{tab:strategies})
and expected $\log(R_{i,t})$ is modelled as:
Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid 
potential confusion like using the term “expected log(R)”.  I also just don't like hats... but that's my problem.
